### 2022-07-19 src/scripts/tree.tex
updates

% Adapted from https://tex.stackexchange.com/a/328890
\documentclass[border=5pt]{standalone}
\usepackage{forest}
\usepackage{fontawesome5}
\newcommand\folderOpen{{\color{black!40!white}\faFolderOpen}\hspace{0.4em}}
\newcommand\folderClosed{{\color{black!40!white}\faFolder}\hspace{0.4em}}
\newcommand\file{{\color{black!40!white}\faFile}\hspace{0.7em}}
\begin{document}
\begin{forest}
  for tree={
    font=\ttfamily,
    grow'=0,
    child anchor=west,
    parent anchor=south,
    anchor=west,
    calign=first,
    inner xsep=2pt,
    edge path={
      \noexpand\path [draw, \forestoption{edge}]
      (!u.south west) +(7.5pt,0) |- (.child anchor) \forestoption{edge label};
    },
    before typesetting nodes={
      if n=1
        {insert before={[,phantom]}}
        {}
    },
    fit=band,
    before computing xy={l=15pt},
  }  
[
  \folderOpen
  [
    \folderOpen .github
    [
      \folderClosed workflows {\sffamily \color{gray} --- Configuration files for GitHub Actions workflows}
    ]
  ]
  [
    \folderOpen src
    [
      \folderClosed data {\sffamily \color{gray} --- Programmatically generated or downloaded datasets}
    ]
    [
      \folderClosed scripts {\sffamily \color{gray} --- Scripts and auxiliary code for generating figures}
    ]
    [
      \folderClosed static {\sffamily \color{gray} --- Version-controlled figures (photographs, charts, etc.)}
    ]
    [
      \folderOpen tex
      [
        \folderClosed figures {\sffamily \color{gray} --- Programmatically generated figures}
      ]
      [
        \folderClosed output {\sffamily \color{gray} --- Other programmatically generated output}
      ]
      [
        \file ms.tex {\sffamily \color{gray} --- Main article manuscript}
      ]
    ]
  ]
  [
    \file Snakefile {\sffamily \color{gray} --- Custom \texttt{snakemake} rules}
  ]
  [
    \file environment.yml {\sffamily \color{gray} --- \texttt{conda} environment specifications}
  ]
  [
    \file showyourwork.yml {\sffamily \color{gray} --- Workflow configuration}
  ]
]
\end{forest}
\end{document}